---
title: "Computer Exercise 1"
author: "Tofig Mamedov & William Lindskog"
date: '2021-04-27'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
load("data/portpdata.Rdata)
```

## Computer Assignment 1 - Report

Introduction...

```{r}
portpirie$models$fit1$results
solve(portpirie$models$fit1$results$hessian)
```

#Q5 Confidence Intervals

You can also embed plots, for example:

```{r question 5}
ci(portpirie$models$fit1, type = "parameter")
```
Results seem to agree as both CI are equal. Same methods but alternative R commands.

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

#Q7 Profile Likelihood CI

```{r}
ci(portpirie$models$fit1, type = "parameter", method = "proflik",
   which.par=3,xrange=c(-0.5,0.5))
```
```{r}
ci(portpirie$models$fit1, type = "parameter", method = "proflik",
   which.par=3,xrange=c(-0.5,0.5))
```

#Q8 ML Estimate of 10 and 100 year return levels
```{r}
params <- portpirie$models$fit1$results$par
y_10 <- -log(0.9)
y_100 <- -log(0.99)
z_10 <- params[1] - params[2]/params[3]*(1-y_10^-params[3])
z_100 <- params[1] - params[2]/params[3]*(1-y_100^-params[3])
# As per page 56 in the book. z_p is the return?
# For variance

z10_delt = c(1, -1/params[3]*(1-y_10^-params[3]),params[2]*params[3]^-2*(1-y_10^-params[3])-params[2]/params[3]*y_10^-params[3]*log(y_10) )
var10 <- t(z10_delt)%*%portpirie$models$fit1$results$hessian%*%z10_delt

z100_delt = c(1, -1/params[3]*(1-y_100^-params[3]),params[2]*params[3]^-2*(1-y_100^-params[3])-params[2]/params[3]*y_100^-params[3]*log(y_100) )
var100 <- t(z100_delt)%*%portpirie$models$fit1$results$hessian%*%z100_delt
```

#Q9 Delta Method
Analyze - confidence interval - normal approx

10 year = (4.1884, 4.404)
100 year = (4.3771, 4.9997)

#Q10 Profile Likelyhood
```{r}
ci(portpirie$models$fit1,type="return.level",return.period = 100,method="proflik",nint=100,xrange=c(4,6))
```
10-year return level: 4.296
95 % CI: (4.2051,4.4441)" 
100-year return level: 4.688
95 % CI: (4.4921,5.2595)

# Q11 - Upper Endpoint
The upper endpoint of the distribution is when p goes to infinity, meaning 1/p goes to zero. As per page 56 in the book this return level can be calculated through ML.
```{r}
z0 <- params[1]-(params[2]/params[3])
z0
```
z0 is 7.83

# Q12 - CI for upper bound
Following the method provided in the book we get
```{r}

```

# Q13 Diagnostic Plots
Plots are too small to see anything
# Q14 Older Diagnostic Plots
```{r}
par(mfrow=c(2,2))
plot(portpirie$models$fit1,type='probprob',main='Probability Plot')
plot(portpirie$models$fit1,type='qq',main='Quantile Plot')
plot(portpirie$models$fit1,type='rl',main='Return Level Plot')
plot(portpirie$models$fit1,type='hist',main='Histogram Plot')
```

# Q15 ML of Gumbel Distribution
```{r}
gumbelparams = portpirie$models$fit2$results$par
gumbelparams[1]
gumbelparams[2]
```

# Q16 Covariance Matrix
```{r}
portpirie$models$fit2$results$hessian
```
# Q17 Confidence Interval
Based on the information matrix. Which method is relevant?
```{r}
portpirie$models$fit2$results$hessian
```
# Q18 Gumbel plots

```{r}
par(mfrow=c(2,2))
plot(portpirie$models$fit2,type='probprob',main='Probability Plot')
plot(portpirie$models$fit2,type='qq',main='Quantile Plot')
plot(portpirie$models$fit2,type='rl',main='Return Level Plot')
plot(portpirie$models$fit2,type='hist',main='Histogram Plot')
```

# Q19 Likelihood Ratio Test

