---
title: "Computer Exercise 1"
author: "Tofig Mamedov & William Lindskog"
date: '2021-04-27'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
load("data/portpdata.Rdata")
```

## Computer Assignment 1 - Report

Introduction...

```{r}
portpirie$models$fit1$results
solve(portpirie$models$fit1$results$hessian)
```

#Q5 Confidence Intervals

You can also embed plots, for example:

```{r question 5}
ci(portpirie$models$fit1, type = "parameter")
```
Results seem to agree as both CI are equal. Same methods but alternative R commands.

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

#Q7 Profile Likelihood CI

```{r}
ci(portpirie$models$fit1, type = "parameter", method = "proflik",
   which.par=3,xrange=c(-0.5,0.5))
```
```{r}
ci(portpirie$models$fit1, type = "parameter", method = "proflik",
   which.par=3,xrange=c(-0.5,0.5))
```

#Q8 ML Estimate of 10 and 100 year return levels
```{r}
params <- portpirie$models$fit1$results$par
y_10 <- -log(0.9)
y_100 <- -log(0.99)
z_10 <- params[1] - params[2]/params[3]*(1-y_10^-params[3])
z_100 <- params[1] - params[2]/params[3]*(1-y_100^-params[3])
# As per page 56 in the book. z_p is the return?
# For variance

z10_delt = c(1, -1/params[3]*(1-y_10^-params[3]),params[2]*params[3]^-2*(1-y_10^-params[3])-params[2]/params[3]*y_10^-params[3]*log(y_10) )
var10 <- t(z10_delt)%*%portpirie$models$fit1$results$hessian%*%z10_delt

z100_delt = c(1, -1/params[3]*(1-y_100^-params[3]),params[2]*params[3]^-2*(1-y_100^-params[3])-params[2]/params[3]*y_100^-params[3]*log(y_100) )
var100 <- t(z100_delt)%*%portpirie$models$fit1$results$hessian%*%z100_delt
```

```{r}

```
